---
title: "Binomial Distribution Exercises"
author: "Arin Parsa"
date: "6/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## OpenIntro Statistics 4th edition, Section 4.3 exercises



4.17 Underage drinking, Part I. Data collected by the Substance Abuse and Mental Health Services Administration (SAMSHA) suggests that 69.7% of 18-20 year olds consumed alcoholic beverages in any given year.

(a) Suppose a random sample of ten 18-20 year olds is taken. Is the use of the binomial distribution
appropriate for calculating the probability that exactly six consumed alcoholic beverages? Explain.
(b) Calculate the probability that exactly 6 out of 10 randomly sampled 18- 20 year olds consumed an
alcoholic drink.

```{r}
dbinom(6, 10, 0.697)
```


(c) What is the probability that exactly four out of ten 18-20 year olds have not consumed an alcoholic
beverage?

```{r}
binom <- choose(10, 4)*(0.303^4)*(0.697^6)
dbinom(4, 10, 0.303)
```


(d) What is the probability that at most 2 out of 5 randomly sampled 18-20 year olds have consumed
alcoholic beverages?

```{r}
pbinom(2, 5, 0.697)
```


(e) What is the probability that at least 1 out of 5 randomly sampled 18-20 year olds have consumed alcoholic beverages

```{r}
1 - pbinom(1, 5, 0.697)
pbinom(1, 5, 0.697, lower.tail = FALSE)
```


